### 2021-03-05 paper1/paper1.tex
Use include instead of import
This has the advantage that I can compile one chapter at a time.

% \documentclass{article}
% \usepackage{style}
% \usepackage{times}

@@

\subsection{Results for Demographic Parity on Adult dataset}\label{sssec:demparresults}
\begin{figure}[t]
  \begin{center}
    \centerline{\includegraphics[width=0.98\columnwidth]{./figures/adult_tar_choice.eps}}
    \caption{Accuracy and fairness (demographic parity) for various target choices.
      (a): Adult dataset using race as the sensitive attribute;
      (b): Adult dataset using gender.
      Centre of the cross is the mean; height and width of the box encode half of standard derivation of accuracy and disparate impact.
    }%
    \label{fig:tar}
  \end{center}
  % \vspace*{-0.8cm}
% \caption{Accuracy and fairness for various targets.  w.r.t. accuracy and width of the box encodes standard derivation w.r.t. fairness criterion}
\end{figure}%
%
Following \citet{zafar2017fairnessconstraints} we evaluate demographic parity on the Adult dataset.
Table~\ref{tab:dempar} shows the accuracy and fairness for several algorithms.
In the table, and in the following, we use $\mathit{PR}_{s=i}$ to denote
the observed rate of positive predictions per demographic group $P (\hat{y} = 1 | s =i)$.
Thus, $\mathit{PR}_{s=0} / \mathit{PR}_{s=1}$ is a measure for demographic parity,
where a completely fair model would attain a value of $1.0$.
This measure for demographic parity is also called ``disparate impact''
(see e.g.~\cite{feldman2015certifying,zafar2017fairnesstreatment}).
%
As the results in Table~\ref{tab:dempar} show, FairGP and FairLR are clearly fairer than the baseline GP and LR\@.
We use the mean ($\mathit{PR}_t^{avg}$) for the target acceptance rate.
The difference between fair models and unconstrained models is not as large with \emph{race} as the sensitive attribute,
as the unconstrained models are already quite fair there.
The results of FairGP are characterised by high fairness and high accuracy.
FairLR achieves similar results to FairGP, but with generally slightly lower accuracy but better fairness.
We used the two step procedure of \citet{DonOneBenShaetal18}
to verify that we cannot achieve the same fairness result with just parameter search on LR\@.

In fig.~\ref{fig:tar}, we investigate which choice of target ($\mathit{PR}_t^{avg}$, $\mathit{PR}_t^{min}$ or $\mathit{PR}_t^{max}$)
gives the best result.
We use $\mathit{PR}_t^{avg}$ for all following experiments as this is the fairest choice (cf. section~\ref{sec:dp}).
The fig.~\ref{fig:tar}(a) shows results from Adult dataset with \emph{race} as sensitive attribute
where we have $\mathit{PR}_t^{min}=0.156$, $\mathit{PR}_t^{max}=0.267$ and $\mathit{PR}_t^{avg} =0.211$.
$\mathit{PR}_t^{avg}$ performs best in term of the trade-off.%equal (for \emph{race} as sensitive attribute)
%or better (for \emph{gender}) compared with the two other possibilities.
% The variant that uses $s$ for training and prediction (``use $s$'') performs significantly better here.
%\emph{ZafarAccuracy} can achieve good fairness results at the cost of accuracy.
\begin{figure}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{./figures/adult_parity_scatter_pr_pr.eps}
  \caption{
    Predictions with different target acceptance rates (demographic parity) for 10 repeats.
    (a): $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$ using race as the sensitive attribute;
    (b): $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$ using gender.
  }%
  \label{fig:adult_parity_scatter_pr_pr}
\end{figure}%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{./figures/adult_parity_scatter_acc.eps}
  \caption{
    Predictions with different target acceptance rates (demographic parity) for 10 repeats.
    (a): disparate impact vs accuracy on Adult dataset using race as the sensitive attribute;
    (b): disparate impact vs accuracy using gender.
  }%
  \label{fig:adult_parity_scatter_acc}
\end{figure}%

@@

\subsection{Results for Equality of Opportunity on ProPublica dataset.}\label{sssec:eqoppresults}
%
For equality of opportunity,
we again follow \citet{zafar2017fairnesstreatment} and evaluate the algorithm on the ProPublica dataset.
As we did for demographic parity,
we define a measure of equality of opportunity via the ratio of the true positive rates (TPRs) within the demographic groups.
We use $\mathit{TPR}_{s=i}$ to denote the observed TPR in group $i$: $P(\hat{y}=1|y=1, s=i)$,
and $\mathit{TNR}_{s=i}$ for the observed true negative rate (TNR) in the same manner.
The measure is then given by $\mathit{TPR}_{s=0} / \mathit{TPR}_{s=1}$.
A perfectly fair algorithm would achieve $1.0$ on the measure.%
\begin{figure*}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{./figures/propublica_opp_box_with_agarwal.eps}
%   \includegraphics[width=0.45\textwidth]{propublica-recidivism_race_opp_box.eps}
%   \includegraphics[width=0.45\textwidth]{propublica-recidivism_sex_opp_box.eps}
  \caption{Accuracy and fairness (with respect to \emph{equality of opportunity}) for various methods on ProPublica dataset.
    (a): using race as the sensitive attribute; (b): using gender.
    A completely fair model would achieve a value of 1.0 in the x-axis.
    %The best value for the TPR ratio is $1.0$.
    See fig.~\ref{fig:propublica_opp_scatter_tpr}(a) and (b) on how these choices of PR setting translate to
    $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$. %and TPR vs TNR.
  }%
  \label{fig:propublica_opp_box}
\end{figure*}%
\begin{figure*}[!ht]
  \centering
  \includegraphics[width=0.98\textwidth]{./figures/propublica_opp_scatter_tpr.eps}
  \caption{%
    Fairness measure $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$ (\emph{equality of opportunity}) for different target PRs ($\mathit{PR}_t$).
    (a): on dataset ProPublica recidivism using race as the sensitive attribute;
    (b): using gender.
  }%
  \label{fig:propublica_opp_scatter_tpr}
\end{figure*}%

@@

% \begin{figure*}[ht]
%   \centering
%   \includegraphics[width=0.45\textwidth]{adult_race_tar.eps}
%   \includegraphics[width=0.45\textwidth]{adult_sex_tar.eps}
%   \caption{
%     Accuracy and fairness (with respect to \emph{demographic parity})
%     in the prediction for our model with different target acceptance rates.
%     Top: Adult dataset using race as the sensitive attribute; Bottom: Adult dataset using gender.
%   }
%   \label{fig:diacc_demo}
% \end{figure*}
% \begin{figure*}[!ht]
% \centering
% \includegraphics[width=0.85\textwidth]{./figures/adult_parity_box.pdf}
% %   \includegraphics[width=0.495\textwidth]{adult_race_demo_box.eps}
% %   \includegraphics[width=0.495\textwidth]{adult_sex_demo_box.eps}
%   \caption{
%     Accuracy and fairness (with respect to \emph{demographic parity}) for various methods.
%     Center of the box is the mean, height of the box encodes the standard deviation w.r.t. accuracy
%     and width of the box encodes the standard deviation w.r.t. DIbinary.
%     The best value of DIbinary is 1.
%     (a): Adult dataset using race as the sensitive attribute; (b): Adult dataset using gender.
%   }
%   \label{fig:adult_parity_box}
% \end{figure*}

% \begin{figure*}[t]
%   \includegraphics[width=0.98\textwidth]{./figures/adult_parity_scatter.pdf}
%   \caption{%
%     Predictions with different target acceptance rates (demographic parity) on Adult dataset.
%     \textbf{(a)} and \textbf{(b)}: $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$.
%     \textbf{(c)} and \textbf{(d)}: $\mathit{PR}_{s=0} / \mathit{PR}_{s=1}$ vs accuracy.
%     Left column:  using race as the sensitive attribute;
%     Right column:  using gender.
%     The \emph{base rate} indicates the positive rates of the training data.
%     ``use s'' indicates that $s$ was appended to the input features.
%   }%
%   \label{fig:adult_parity_scatter}
% \end{figure*}
% \begin{figure*}[t]
%   \includegraphics[width=0.98\textwidth]{./figures/eo_together.pdf}
%   \caption{%
%     Predictions with different target true positive rates ($\mathit{TPR}_t$; equality of opportunity) on ProPublica dataset.
%     Our results were obtained with the Logistic Regression model.
%     \textbf{(a)} and \textbf{(b)}: $\mathit{TPR}_{s=0} / \mathit{TPR}_{s=1}$ vs accuracy.
%     \textbf{(c)} and \textbf{(d)}: $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$.
%     Left column:  using race as the sensitive attribute;
%     Right column:  using gender.
%     $s$ is \emph{not} used as an input feature.
%   }%
%   \label{fig:propublica_opp_scatter}
% \end{figure*}
%
%

@@

% Figures \ref{fig:odds_tpr_opp} and \ref{fig:odds_opp} show the results from experiments
% where the target TPR rate was set manually
% instead of being estimated from training and evaluating on the pre-validation and pre-training set.
% \begin{figure*}[!ht]
%   \centering
%   \includegraphics[width=0.98\textwidth]{./figures/propublica_opp_scatter_tnr.pdf}
%   \caption{%
%     $\mathit{TPR}$ vs $\mathit{TNR}$ for different target TPRs ($\mathit{TPR}_t$).
%     (a): on dataset ProPublica recidivism using race as the sensitive attribute;
%     (b): using gender.
%   }%
%   \label{fig:propublica_opp_scatter_tnr}
% \end{figure*}

@@

  Let $d_s$ be the new decision boundary: $d_s=\tfrac{1}{2}\cdot m_s + b_s$. There are two possibilities to consider here:
  either $\tfrac{1}{2}<d_s$ or $\tfrac{1}{2}>d_s$ (for $d_s=\tfrac{1}{2}$, the decision boundaries are the same and nothing has to be shown).
  The problem, $y\neq\bar{y}$, appears then exactly when the value of $\bar{\eta}$ is between the two boundaries:
  \begin{align}
    \text{if}~d_s>\tfrac{1}{2}\text{:}\quad d_s>\bar{\eta}(x, s)>\tfrac{1}{2} \\
    \text{if}~d_s<\tfrac{1}{2}\text{:}\quad d_s<\bar{\eta}(x, s)<\tfrac{1}{2}
  \end{align}
  Expressing this in terms of $\eta$ and simplifying leads to
  (if $m_s$ is negative, then the two cases are swapped, but we still get both inequalities):
  \begin{align}
    \text{if}~d_s>\tfrac{1}{2}\text{:}\quad\frac{1}{2}>\eta(x,s)>\frac{1-2b_s}{2m_s} \\
    \text{if}~d_s<\tfrac{1}{2}\text{:}\quad\frac{1}{2}<\eta(x,s)<\frac{1-2b_s}{2m_s}
  \end{align}
  This can be summarized as
  \begin{align}
    \left|\eta(x,s)-\frac{1}{2}\right|<\left|\frac{1}{2}-\frac{1-2b_s}{2m_s}\right|~.
  \end{align}
  Let $t_s$ denote the term on the right side of this inequality (i.e. the ``threshold'' that determines whether $y=\bar{y}$ or not). Then
  \begin{align}
    t_s=\left|\frac{1}{2} -\frac{1-2b_s}{2m_s}\right|=\left|\frac{m_s+2b_s-1}{2m_s}\right|~.
  \end{align}
  So, we have: $\left|\eta(x, s) - \frac{1}{2}\right| < t_s =\left|\tfrac{m_s+2b_s-1}{2m_s}\right|$.
  This leads directly to the statement we wanted to prove:
  \begin{align}
    P(y\neq\bar{y}|s)=P\left(\left|\eta(x, s)-\frac{1}{2}\right| <t_s\right)~.
  \end{align}
\end{proof}

\subsection{Finding minimal $t_s$}
We express $t_s$ in terms of $\mathit{PR}_b^s$ and $\mathit{PR}_t$.
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr2}
\end{align}
Without loss of generality, we assume $\mathit{PR}_b^0<\mathit{PR}_b^1$.
As mentioned in the main text, 
$\mathit{PR}_t$ should be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$ to minimize both $t_s$.
If that is the case, then we get
\begin{align}
  t_{s=0} &= \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^0}{1 - \mathit{PR}_t}\\
  t_{s=1} &= \frac{1}{2}\frac{\mathit{PR}_b^1 - \mathit{PR}_t}{\mathit{PR}_t}~.
\end{align}
If we further assume $\mathit{PR}_b^1<\tfrac{1}{2}$,
then we also have $\mathit{PR}_t<\tfrac{1}{2}$ and thus $\mathit{PR}_t<1-\mathit{PR}_t$.
This implies that the denominator of $t_{s=1}$ is smaller and that, in turn, $t_{s=1}$ grows faster.
This faster growth means that when minimizing $t_{s=0} + t_{s=1}$, we have to concentrate on $t_{s=1}$.
The minimum is then such that $t_{s=1}$ is 0, i.e.\ $\mathit{PR}_t=\mathit{PR}_b^1$.

@@

  For $\hat{r}_a > r_a$, the roles of $\mathcal{C}^0$ and $\mathcal{C}^1$ are reversed
  and thus, the signs in the equation are inverted:
  \begin{align}
    \left|\mathcal{A}^\mathit{optimal}\right| = N  - (\left|\mathcal{Y}^1\right| - \left|\mathcal{C}^1\right|)~.
  \end{align}
  This proves the claim.
\end{proof}
\begin{corollary}
  Given a dataset that consists of two subsets $\mathcal{S}_0$
  and $\mathcal{S}_1$ ($\mathcal{D} = \mathcal{S}_0 \cup \mathcal{S}_1$)
  where $p$ is the ratio of $|\mathcal{S}_0|$ to $|\mathcal{D}|$
  and given corresponding acceptance rates $r^0_a$ and $r^1_a$
  and predictions with target rates $\hat{r}^0_a$ and $\hat{r}^1_a$,
  the accuracy is limited by
  \begin{align}
    \mathit{acc} \leq 1 - p\cdot\left| \hat{r}^0_a -r^0_a \right| - (1-p)\cdot\left| \hat{r}^1_a -r^1_a \right|~.
  \end{align}
\end{corollary}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{./figures/plot.pdf}
  \caption{Achievable accuracy for different target values.}%
  \label{fig:accvstarget}
\end{figure}
\begin{example}
  We consider the case where $S_0$
  (which could for example be all data points for female individuals) makes up 30\% of the dataset; so $p = 0.3$.
  Further, we say that for $S_0$ we have an acceptance rate of 10\% ($r^0_a = 0.1$)
  and for $S_1$, 50\% ($r^1_a = 0.5$).
  If we then set both target rates to the same value $t$ ($\hat{r}^0_a=\hat{r}^1_a=t$), with $t = 0.3$,
  then the highest accuracy that can be achieved is $0.8$ or 80\%.

@@

\subsection{Illustration of restrictions on PR}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/label_shift.pdf}
  \caption{%
    Illustration of demographic parity with target labels.
    In the situation in the upper part, $P(\bar{y}=1|y=1)$ cannot be set to 1,
    because there are more samples with $y=1$ than there are $\bar{y}=1$.
    In the situation in the lower part, $P(\bar{y}=1|y=1)$ can be set to 1.
  }%
  \label{fig:shift}
\end{figure}%
We start by setting a target rate $r_t$:
\begin{align}
  P(\bar{y}=1|s=0) \overset{!}{=} r_t \quad\text{and}\quad P(\bar{y}=1|s=1) \overset{!}{=} r_t
\end{align}

@@

For $P(y|s=s\prime)$ we will put in the value at which we want our constraint to hold.
We use $r_b^j$ to denote the base rate $P(y=1|s=j)$ which we estimate from the training set.
Plugging this in, we are left with
\begin{align}
  r_t &= P(\bar{y}=1|y=0,s=0) \cdot (1-r_b^0) +P(\bar{y}=1|y=1,s=0) \cdot r_b^0 \\
  r_t &= P(\bar{y}=1|y=0,s=1) \cdot (1-r_b^1) +P(\bar{y}=1|y=1,s=1) \cdot r_b^1~.
\end{align}
This is a system of linear equations with two equations and four free variables.
There is thus still considerable freedom in how we want our constraint to be realized.
The freedom that we have here concerns how strongly the accuracy will be affected.

If we set $P(\bar{y}=1|y=1,s)$ to 0.5,
then we express the fact that a train label of 1 only implies a target label of 1 in 50\% of the cases.
In order to minimize the effect on accuracy,
we make $P(\bar{y}=1|y=1,s)$ as high as possible and $P(\bar{y}=1|y=0,s)$ as low as possible.

>>>>

\chapter{Paper 1: Tuning Fairness by Balancing Target Labels}\label{ch:paper1}
% \documentclass{article}
% \usepackage{style}
% \usepackage{times}

@@

\subsection{Results for Demographic Parity on Adult dataset}\label{sssec:demparresults}
\begin{figure}[t]
  \begin{center}
    \centerline{\includegraphics[width=0.98\columnwidth]{paper1/figures/adult_tar_choice.eps}}
    \caption{Accuracy and fairness (demographic parity) for various target choices.
      (a): Adult dataset using race as the sensitive attribute;
      (b): Adult dataset using gender.
      Centre of the cross is the mean; height and width of the box encode half of standard derivation of accuracy and disparate impact.
    }%
    \label{fig:tar}
  \end{center}
  % \vspace*{-0.8cm}
% \caption{Accuracy and fairness for various targets.  w.r.t. accuracy and width of the box encodes standard derivation w.r.t. fairness criterion}
\end{figure}%
%
Following \citet{zafar2017fairnessconstraints} we evaluate demographic parity on the Adult dataset.
Table~\ref{tab:dempar} shows the accuracy and fairness for several algorithms.
In the table, and in the following, we use $\mathit{PR}_{s=i}$ to denote
the observed rate of positive predictions per demographic group $P (\hat{y} = 1 | s =i)$.
Thus, $\mathit{PR}_{s=0} / \mathit{PR}_{s=1}$ is a measure for demographic parity,
where a completely fair model would attain a value of $1.0$.
This measure for demographic parity is also called ``disparate impact''
(see e.g.~\cite{feldman2015certifying,zafar2017fairnesstreatment}).
%
As the results in Table~\ref{tab:dempar} show, FairGP and FairLR are clearly fairer than the baseline GP and LR\@.
We use the mean ($\mathit{PR}_t^{avg}$) for the target acceptance rate.
The difference between fair models and unconstrained models is not as large with \emph{race} as the sensitive attribute,
as the unconstrained models are already quite fair there.
The results of FairGP are characterised by high fairness and high accuracy.
FairLR achieves similar results to FairGP, but with generally slightly lower accuracy but better fairness.
We used the two step procedure of \citet{DonOneBenShaetal18}
to verify that we cannot achieve the same fairness result with just parameter search on LR\@.

In fig.~\ref{fig:tar}, we investigate which choice of target ($\mathit{PR}_t^{avg}$, $\mathit{PR}_t^{min}$ or $\mathit{PR}_t^{max}$)
gives the best result.
We use $\mathit{PR}_t^{avg}$ for all following experiments as this is the fairest choice (cf. section~\ref{sec:dp}).
The fig.~\ref{fig:tar}(a) shows results from Adult dataset with \emph{race} as sensitive attribute
where we have $\mathit{PR}_t^{min}=0.156$, $\mathit{PR}_t^{max}=0.267$ and $\mathit{PR}_t^{avg} =0.211$.
$\mathit{PR}_t^{avg}$ performs best in term of the trade-off.%equal (for \emph{race} as sensitive attribute)
%or better (for \emph{gender}) compared with the two other possibilities.
% The variant that uses $s$ for training and prediction (``use $s$'') performs significantly better here.
%\emph{ZafarAccuracy} can achieve good fairness results at the cost of accuracy.
\begin{figure}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{paper1/figures/adult_parity_scatter_pr_pr.eps}
  \caption{
    Predictions with different target acceptance rates (demographic parity) for 10 repeats.
    (a): $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$ using race as the sensitive attribute;
    (b): $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$ using gender.
  }%
  \label{fig:adult_parity_scatter_pr_pr}
\end{figure}%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{paper1/figures/adult_parity_scatter_acc.eps}
  \caption{
    Predictions with different target acceptance rates (demographic parity) for 10 repeats.
    (a): disparate impact vs accuracy on Adult dataset using race as the sensitive attribute;
    (b): disparate impact vs accuracy using gender.
  }%
  \label{fig:adult_parity_scatter_acc}
\end{figure}%

@@

\subsection{Results for Equality of Opportunity on ProPublica dataset.}\label{sssec:eqoppresults}
%
For equality of opportunity,
we again follow \citet{zafar2017fairnesstreatment} and evaluate the algorithm on the ProPublica dataset.
As we did for demographic parity,
we define a measure of equality of opportunity via the ratio of the true positive rates (TPRs) within the demographic groups.
We use $\mathit{TPR}_{s=i}$ to denote the observed TPR in group $i$: $P(\hat{y}=1|y=1, s=i)$,
and $\mathit{TNR}_{s=i}$ for the observed true negative rate (TNR) in the same manner.
The measure is then given by $\mathit{TPR}_{s=0} / \mathit{TPR}_{s=1}$.
A perfectly fair algorithm would achieve $1.0$ on the measure.%
\begin{figure*}[t]
  \centering
  \includegraphics[width=0.98\textwidth]{paper1/figures/propublica_opp_box_with_agarwal.eps}
%   \includegraphics[width=0.45\textwidth]{propublica-recidivism_race_opp_box.eps}
%   \includegraphics[width=0.45\textwidth]{propublica-recidivism_sex_opp_box.eps}
  \caption{Accuracy and fairness (with respect to \emph{equality of opportunity}) for various methods on ProPublica dataset.
    (a): using race as the sensitive attribute; (b): using gender.
    A completely fair model would achieve a value of 1.0 in the x-axis.
    %The best value for the TPR ratio is $1.0$.
    See fig.~\ref{fig:propublica_opp_scatter_tpr}(a) and (b) on how these choices of PR setting translate to
    $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$. %and TPR vs TNR.
  }%
  \label{fig:propublica_opp_box}
\end{figure*}%
\begin{figure*}[!ht]
  \centering
  \includegraphics[width=0.98\textwidth]{paper1/figures/propublica_opp_scatter_tpr.eps}
  \caption{%
    Fairness measure $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$ (\emph{equality of opportunity}) for different target PRs ($\mathit{PR}_t$).
    (a): on dataset ProPublica recidivism using race as the sensitive attribute;
    (b): using gender.
  }%
  \label{fig:propublica_opp_scatter_tpr}
\end{figure*}%

@@

% \begin{figure*}[ht]
%   \centering
%   \includegraphics[width=0.45\textwidth]{adult_race_tar.eps}
%   \includegraphics[width=0.45\textwidth]{adult_sex_tar.eps}
%   \caption{
%     Accuracy and fairness (with respect to \emph{demographic parity})
%     in the prediction for our model with different target acceptance rates.
%     Top: Adult dataset using race as the sensitive attribute; Bottom: Adult dataset using gender.
%   }
%   \label{fig:diacc_demo}
% \end{figure*}
% \begin{figure*}[!ht]
% \centering
% \includegraphics[width=0.85\textwidth]{paper1/figures/adult_parity_box.pdf}
% %   \includegraphics[width=0.495\textwidth]{adult_race_demo_box.eps}
% %   \includegraphics[width=0.495\textwidth]{adult_sex_demo_box.eps}
%   \caption{
%     Accuracy and fairness (with respect to \emph{demographic parity}) for various methods.
%     Center of the box is the mean, height of the box encodes the standard deviation w.r.t. accuracy
%     and width of the box encodes the standard deviation w.r.t. DIbinary.
%     The best value of DIbinary is 1.
%     (a): Adult dataset using race as the sensitive attribute; (b): Adult dataset using gender.
%   }
%   \label{fig:adult_parity_box}
% \end{figure*}

% \begin{figure*}[t]
%   \includegraphics[width=0.98\textwidth]{paper1/figures/adult_parity_scatter.pdf}
%   \caption{%
%     Predictions with different target acceptance rates (demographic parity) on Adult dataset.
%     \textbf{(a)} and \textbf{(b)}: $\mathit{PR}_{s=0}$ vs $\mathit{PR}_{s=1}$.
%     \textbf{(c)} and \textbf{(d)}: $\mathit{PR}_{s=0} / \mathit{PR}_{s=1}$ vs accuracy.
%     Left column:  using race as the sensitive attribute;
%     Right column:  using gender.
%     The \emph{base rate} indicates the positive rates of the training data.
%     ``use s'' indicates that $s$ was appended to the input features.
%   }%
%   \label{fig:adult_parity_scatter}
% \end{figure*}
% \begin{figure*}[t]
%   \includegraphics[width=0.98\textwidth]{paper1/figures/eo_together.pdf}
%   \caption{%
%     Predictions with different target true positive rates ($\mathit{TPR}_t$; equality of opportunity) on ProPublica dataset.
%     Our results were obtained with the Logistic Regression model.
%     \textbf{(a)} and \textbf{(b)}: $\mathit{TPR}_{s=0} / \mathit{TPR}_{s=1}$ vs accuracy.
%     \textbf{(c)} and \textbf{(d)}: $\mathit{TPR}_{s=0}$ vs $\mathit{TPR}_{s=1}$.
%     Left column:  using race as the sensitive attribute;
%     Right column:  using gender.
%     $s$ is \emph{not} used as an input feature.
%   }%
%   \label{fig:propublica_opp_scatter}
% \end{figure*}
%
%

@@

% Figures \ref{fig:odds_tpr_opp} and \ref{fig:odds_opp} show the results from experiments
% where the target TPR rate was set manually
% instead of being estimated from training and evaluating on the pre-validation and pre-training set.
% \begin{figure*}[!ht]
%   \centering
%   \includegraphics[width=0.98\textwidth]{paper1/figures/propublica_opp_scatter_tnr.pdf}
%   \caption{%
%     $\mathit{TPR}$ vs $\mathit{TNR}$ for different target TPRs ($\mathit{TPR}_t$).
%     (a): on dataset ProPublica recidivism using race as the sensitive attribute;
%     (b): using gender.
%   }%
%   \label{fig:propublica_opp_scatter_tnr}
% \end{figure*}

@@

  Let $d_s$ be the new decision boundary: $d_s=\tfrac{1}{2}\cdot m_s + b_s$. There are two possibilities to consider here:
  either $\tfrac{1}{2}<d_s$ or $\tfrac{1}{2}>d_s$ (for $d_s=\tfrac{1}{2}$, the decision boundaries are the same and nothing has to be shown).
  The problem, $y\neq\bar{y}$, appears then exactly when the value of $\bar{\eta}$ is between the two boundaries:
  \begin{align}
    \text{if}~d_s>\tfrac{1}{2}\text{:}\quad d_s>\bar{\eta}(x, s)>\tfrac{1}{2} \\
    \text{if}~d_s<\tfrac{1}{2}\text{:}\quad d_s<\bar{\eta}(x, s)<\tfrac{1}{2}
  \end{align}
  Expressing this in terms of $\eta$ and simplifying leads to
  (if $m_s$ is negative, then the two cases are swapped, but we still get both inequalities):
  \begin{align}
    \text{if}~d_s>\tfrac{1}{2}\text{:}\quad\frac{1}{2}>\eta(x,s)>\frac{1-2b_s}{2m_s} \\
    \text{if}~d_s<\tfrac{1}{2}\text{:}\quad\frac{1}{2}<\eta(x,s)<\frac{1-2b_s}{2m_s}
  \end{align}
  This can be summarised as
  \begin{align}
    \left|\eta(x,s)-\frac{1}{2}\right|<\left|\frac{1}{2}-\frac{1-2b_s}{2m_s}\right|~.
  \end{align}
  Let $t_s$ denote the term on the right side of this inequality (i.e. the ``threshold'' that determines whether $y=\bar{y}$ or not). Then
  \begin{align}
    t_s=\left|\frac{1}{2} -\frac{1-2b_s}{2m_s}\right|=\left|\frac{m_s+2b_s-1}{2m_s}\right|~.
  \end{align}
  So, we have: $\left|\eta(x, s) - \frac{1}{2}\right| < t_s =\left|\tfrac{m_s+2b_s-1}{2m_s}\right|$.
  This leads directly to the statement we wanted to prove:
  \begin{align}
    P(y\neq\bar{y}|s)=P\left(\left|\eta(x, s)-\frac{1}{2}\right| <t_s\right)~.
  \end{align}
\end{proof}

\subsection{Finding minimal $t_s$}
We express $t_s$ in terms of $\mathit{PR}_b^s$ and $\mathit{PR}_t$.
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr2}
\end{align}
Without loss of generality, we assume $\mathit{PR}_b^0<\mathit{PR}_b^1$.
As mentioned in the main text, 
$\mathit{PR}_t$ should be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$ to minimise both $t_s$.
If that is the case, then we get
\begin{align}
  t_{s=0} &= \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^0}{1 - \mathit{PR}_t}\\
  t_{s=1} &= \frac{1}{2}\frac{\mathit{PR}_b^1 - \mathit{PR}_t}{\mathit{PR}_t}~.
\end{align}
If we further assume $\mathit{PR}_b^1<\tfrac{1}{2}$,
then we also have $\mathit{PR}_t<\tfrac{1}{2}$ and thus $\mathit{PR}_t<1-\mathit{PR}_t$.
This implies that the denominator of $t_{s=1}$ is smaller and that, in turn, $t_{s=1}$ grows faster.
This faster growth means that when minimizing $t_{s=0} + t_{s=1}$, we have to concentrate on $t_{s=1}$.
The minimum is then such that $t_{s=1}$ is 0, i.e.\ $\mathit{PR}_t=\mathit{PR}_b^1$.

@@

  For $\hat{r}_a > r_a$, the roles of $\mathcal{C}^0$ and $\mathcal{C}^1$ are reversed
  and thus, the signs in the equation are inverted:
  \begin{align}
    \left|\mathcal{A}^\mathit{optimal}\right| = N  - (\left|\mathcal{Y}^1\right| - \left|\mathcal{C}^1\right|)~.
  \end{align}
  This proves the claim.
\end{proof}
\begin{corollary}
  Given a dataset that consists of two subsets $\mathcal{S}_0$
  and $\mathcal{S}_1$ ($\mathcal{D} = \mathcal{S}_0 \cup \mathcal{S}_1$)
  where $p$ is the ratio of $|\mathcal{S}_0|$ to $|\mathcal{D}|$
  and given corresponding acceptance rates $r^0_a$ and $r^1_a$
  and predictions with target rates $\hat{r}^0_a$ and $\hat{r}^1_a$,
  the accuracy is limited by
  \begin{align}
    \mathit{acc} \leq 1 - p\cdot\left| \hat{r}^0_a -r^0_a \right| - (1-p)\cdot\left| \hat{r}^1_a -r^1_a \right|~.
  \end{align}
\end{corollary}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{paper1/figures/plot.pdf}
  \caption{Achievable accuracy for different target values.}%
  \label{fig:accvstarget}
\end{figure}
\begin{example}
  We consider the case where $S_0$
  (which could for example be all data points for female individuals) makes up 30\% of the dataset; so $p = 0.3$.
  Further, we say that for $S_0$ we have an acceptance rate of 10\% ($r^0_a = 0.1$)
  and for $S_1$, 50\% ($r^1_a = 0.5$).
  If we then set both target rates to the same value $t$ ($\hat{r}^0_a=\hat{r}^1_a=t$), with $t = 0.3$,
  then the highest accuracy that can be achieved is $0.8$ or 80\%.

@@

\subsection{Illustration of restrictions on PR}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\textwidth]{paper1/figures/label_shift.pdf}
  \caption{%
    Illustration of demographic parity with target labels.
    In the situation in the upper part, $P(\bar{y}=1|y=1)$ cannot be set to 1,
    because there are more samples with $y=1$ than there are $\bar{y}=1$.
    In the situation in the lower part, $P(\bar{y}=1|y=1)$ can be set to 1.
  }%
  \label{fig:shift}
\end{figure}%
We start by setting a target rate $r_t$:
\begin{align}
  P(\bar{y}=1|s=0) \overset{!}{=} r_t \quad\text{and}\quad P(\bar{y}=1|s=1) \overset{!}{=} r_t
\end{align}

@@

For $P(y|s=s\prime)$ we will put in the value at which we want our constraint to hold.
We use $r_b^j$ to denote the base rate $P(y=1|s=j)$ which we estimate from the training set.
Plugging this in, we are left with
\begin{align}
  r_t &= P(\bar{y}=1|y=0,s=0) \cdot (1-r_b^0) +P(\bar{y}=1|y=1,s=0) \cdot r_b^0 \\
  r_t &= P(\bar{y}=1|y=0,s=1) \cdot (1-r_b^1) +P(\bar{y}=1|y=1,s=1) \cdot r_b^1~.
\end{align}
This is a system of linear equations with two equations and four free variables.
There is thus still considerable freedom in how we want our constraint to be realised.
The freedom that we have here concerns how strongly the accuracy will be affected.

If we set $P(\bar{y}=1|y=1,s)$ to 0.5,
then we express the fact that a train label of 1 only implies a target label of 1 in 50\% of the cases.
In order to minimise the effect on accuracy,
we make $P(\bar{y}=1|y=1,s)$ as high as possible and $P(\bar{y}=1|y=0,s)$ as low as possible.
